\bf A Unitary RG-based Auxiliary Model Approach to Strongly-Correlated Electronic Systems

### 2023-05-29 TilingProjectDraft/tilingProject.tex
--no-message

	\rho_{\vec k,\sigma} = \sum_{n_{\vec k,\sigma}=0,1}\braket{n_{k,\sigma} | \Psi_\text{gs}}\braket{\Psi_\text{gs} | n_{\vec k,\sigma}} = \sum_{n_{\vec k,\sigma}=0,1}|\braket{\Psi_\text{gs} | n_{\vec k,\sigma}}|^2

@@

where \(\ket{n_{\vec k,\sigma}}\) represents a state in Fock space with \(n_{\vec k,\sigma}\) number of particles in the single particle state with momentum \(\vec k\) and spin \(\sigma\). We now Fourier transform the overlap and retain terms up to the next-nearest neighbour:

@@

							      & + 2\Xi_{\vec k}\braket{\Psi_\text{gs} | n_{\vec R_0,\sigma}}\braket{n_{\vec R_0 + \vec a^\prime,\sigma} | \Psi_\text{gs}} 

@@

where \(\vec a\) and \(\vec a^\prime\) are nearest and next-nearest neighbour sites to \(\vec R_0\), and \(\xi_{\vec k}\) and \(\Xi_{\vec k}\) have been defined above. Following the eqs.~\ref{local_correlation} through \ref{nnn_correlation}, we can identify \(\hat O_1^\dagger(\vec R) = \hat O_2(\vec R) = \ket{n_{\vec R_0,\sigma}}\bra{n_{\vec R,\sigma}}\), so that we have, for example, \(\braket{\Psi_\text{gs} | n_{\vec R_0,\sigma}}\braket{n_{\vec R_0 + \vec a,\sigma} | \Psi_\text{gs}} = \braket{\hat O_1(\vec R_0) \hat O_2(\vec R_0 + \vec a)}\). Then, using eqs.~\ref{local_correlation} through \ref{nnn_correlation}, the Fourier transform becomes:

@@

	|\braket{\Psi_\text{gs} | n_{\vec k,\sigma}}|^2 \simeq& \frac{1 + \xi_{\vec k}^2 + \Xi_{\vec k}^2}{3\lambda_{\vec k_0}^2}\left[\braket{\psi_\text{aux} | n_{d,\sigma}}\braket{n_{d,\sigma} | \psi_\text{aux}} + \frac{1}{\mathcal{W}}\braket{\psi_\text{aux} | n_{z,\sigma}}\braket{n_{z,\sigma} | \psi_\text{aux}} + \frac{1}{\mathcal{W}}\braket{\psi_\text{aux} | n_{1,\sigma}}\braket{n_{1,\sigma} | \psi_\text{aux}}\right]\\
							      &+ \frac{2\xi_{\vec k}\left(1 + \Xi_{\vec k}\right)}{4\lambda_{\vec k_0}^2 \mathcal{W}}\left[\braket{\psi_\text{aux} \left(\ket{n_{d,\sigma}}\bra{n_{z,\sigma}} + \ket{n_{z,\sigma}}\bra{n_{1,\sigma}} + \ket{n_{z,\sigma}}\bra{n_{d,\sigma}} + \ket{n_{1,\sigma}}\bra{n_{z,\sigma}}\right) \psi_\text{aux}}\right] \\
							      &+ \frac{2\Xi_{\vec k}}{2\lambda_{\vec k_0}^2 \mathcal{W}}\left[\braket{\psi_\text{aux} \left(\ket{n_{d,\sigma}}\bra{n_{1,\sigma}} + \ket{n_{1,\sigma}}\bra{n_{d,\sigma}}\right) \psi_\text{aux}}\right]

@@

The \(k-\)space reduced density matrix takes the form

@@

	\rho_{\vec k,\sigma} =& \sum_{n_{(d,z,1),\sigma}=0,1}\frac{1 + \xi_{\vec k}^2 + \Xi_{\vec k}^2}{3\lambda_{\vec k_0}^2}\left[\bar\psi^\dagger_{n_{d\sigma}}\bar\psi_{n_{d\sigma}} + \frac{1}{\mathcal{W}}\bar\psi^\dagger_{n_{z\sigma}}\bar\psi_{n_{z\sigma}} + \frac{1}{\mathcal{W}}\bar\psi^\dagger_{n_{1\sigma}}\bar\psi_{n_{1\sigma}}\right] + \sum_{n_{(d,1),\sigma}=0,1}\frac{\Xi_{\vec k}}{\lambda_{\vec k_0}^2 \mathcal{W}}\left[\bar\psi^\dagger_{n_{1\sigma}}\bar\psi_{n_{d\sigma}} + \bar\psi^\dagger_{n_{d\sigma}}\bar\psi_{n_{1\sigma}}\right]\\
							      &+ \sum_{n_{(d,z,1),\sigma}=0,1}\frac{\xi_{\vec k}\left(1 + \Xi_{\vec k}\right)}{2\lambda_{\vec k_0}^2 \mathcal{W}}\left[\bar\psi^\dagger_{n_{d\sigma}}\bar\psi_{n_{z\sigma}} + \bar\psi^\dagger_{n_{z\sigma}}\bar\psi_{n_{d\sigma}} + \bar\psi^\dagger_{n_{1\sigma}}\bar\psi_{n_{z\sigma}} + \bar\psi^\dagger_{n_{z\sigma}}\bar\psi_{n_{1\sigma}}\right]

@@

where \(\bar\psi_{n_{d\sigma}} = \braket{n_{d,\sigma} | \psi_\text{aux}}\) and similar definitions exist for \(\bar\psi_{n_{z\sigma}}\) and \(\bar\psi_{n_{1\sigma}}\). The von-Neumann entropy \(S_\text{EE}(\vec k)\) can then be calculated by taking the trace of the operator \(\rho_{\vec k,\sigma}\ln \rho_{\vec k,\sigma}\).

>>>>

	\rho\left(\vec k\sigma\right) = \sum_{n_{\vec k,\sigma}=0,1}\braket{n_{k,\sigma} | \Psi_\text{gs}}\braket{\Psi_\text{gs} | n_{\vec k,\sigma}} = \sum_{n_{\vec k,\sigma}=0,1}|\braket{\Psi_\text{gs} | n_{\vec k,\sigma}}|^2

@@

where \(\ket{n_{\vec k,\sigma}}\) represents a state in Fock space with \(n_{\vec k,\sigma}\) number of particles in the single particle state with momentum \(\vec k\) and spin \(\sigma\). We now Fourier transform the overlap and retain terms up to the next-nearest neighbour, by expanding \(\ket{n_{\vec k\sigma}}\) as \(\ket{n_{\vec R_0,\sigma}} + \xi_{\vec k}\ket{n_{\vec R_0 + \vec a,\sigma}} + \Xi_{\vec k}\ket{n_{\vec R_0 + \vec a^\prime,\sigma}}\), where \(\vec a\) and \(\vec a^\prime\) are nearest and next-nearest neighbour sites to \(\vec R_0\), and \(\xi_{\vec k}\) and \(\Xi_{\vec k}\) have been defined above. The overlap takes the form

@@

	 & + 2\Xi_{\vec k}\braket{\Psi_\text{gs} | n_{\vec R_0,\sigma}}\braket{n_{\vec R_0 + \vec a^\prime,\sigma} | \Psi_\text{gs}}~.\label{overlap_ftransform}

@@

Following the eqs.~\ref{local_correlation} through \ref{nnn_correlation}, we can identify \(\hat O_1^\dagger(\vec R) = \hat O_2(\vec R) = \ket{n_{\vec R_0,\sigma}}\bra{n_{\vec R,\sigma}}\), so that we have, for example, \(\braket{\Psi_\text{gs} | n_{\vec R_0,\sigma}}\braket{n_{\vec R_0 + \vec a,\sigma} | \Psi_\text{gs}} = \braket{\hat O_1(\vec R_0) \hat O_2(\vec R_0 + \vec a)}\). Then, using eqs.~\ref{local_correlation} through \ref{nnn_correlation}, the Fourier transform becomes:

@@

|\braket{\Psi_\text{gs} | n_{\vec k,\sigma}}|^2 \simeq& \frac{1 + \xi_{\vec k}^2 + \Xi_{\vec k}^2}{3\lambda_{\vec k_0}^2}\left[\braket{n_{d,\sigma} | \rho_\text{aux} | n_{d,\sigma}} + \frac{1}{\mathcal{W}}\braket{n_{z,\sigma} | \rho_\text{aux} | n_{z,\sigma}} + \frac{1}{\mathcal{W}}\braket{n_{1,\sigma} | \rho_\text{aux} | n_{1,\sigma}}\right]\\
							      &+ \frac{2\xi_{\vec k}\left(1 + \Xi_{\vec k}\right)}{4\lambda_{\vec k_0}^2 \mathcal{W}}\left[\braket{n_{d,\sigma} | \rho_\text{aux} | n_{z,\sigma}} + \braket{n_{z,\sigma} | \rho_\text{aux} | n_{d,\sigma}} + \braket{n_{z,\sigma} | \rho_\text{aux} | n_{1,\sigma}} + \braket{n_{1,\sigma} | \rho_\text{aux} | n_{z,\sigma}} \right] \\
							      &+ \frac{2\Xi_{\vec k}}{2\lambda_{\vec k_0}^2 \mathcal{W}}\left[\braket{n_{1,\sigma} | \rho_\text{aux} | n_{d,\sigma}} + \braket{n_{d,\sigma} | \rho_\text{aux} | n_{1,\sigma}}\right]

@@

where \(\rho_\text{aux} \equiv \ket{\psi_\text{aux}}\bra{\psi_\text{aux}}\) is the density matrix corresponding to the auxiliary model ground state. The \(k-\)space reduced density matrix takes the form

@@

	\rho\left(\vec k\sigma\right) =& \sum_{n_{(d,z,1),\sigma}=0,1}\frac{1 + \xi_{\vec k}^2 + \Xi_{\vec k}^2}{3\lambda_{\vec k_0}^2}\left[\braket{n_{d,\sigma} | \rho_\text{aux} | n_{d,\sigma}} + \frac{1}{\mathcal{W}}\braket{n_{z,\sigma} | \rho_\text{aux} | n_{z,\sigma}} + \frac{1}{\mathcal{W}}\braket{n_{1,\sigma} | \rho_\text{aux} | n_{1,\sigma}}\right] \\
				       &+ \sum_{n_{(d,z,1),\sigma}=0,1}\frac{\xi_{\vec k}\left(1 + \Xi_{\vec k}\right)}{2\lambda_{\vec k_0}^2 \mathcal{W}}\left[\braket{n_{d,\sigma} | \rho_\text{aux} | n_{z,\sigma}} + \braket{n_{z,\sigma} | \rho_\text{aux} | n_{d,\sigma}} + \braket{n_{z,\sigma} | \rho_\text{aux} | n_{1,\sigma}} + \braket{n_{1,\sigma} | \rho_\text{aux} | n_{z,\sigma}} \right]\label{density_matrix_ksapce}\\
				       &+ \sum_{n_{(d,1),\sigma}=0,1}\frac{\Xi_{\vec k}}{\lambda_{\vec k_0}^2 \mathcal{W}}\left[\braket{n_{1,\sigma} | \rho_\text{aux} | n_{d,\sigma}} + \braket{n_{d,\sigma} | \rho_\text{aux} | n_{1,\sigma}}\right]~.

@@

The von-Neumann entropy \(S_\text{EE}(\vec k)\) can then be calculated by taking the trace of the operator \(\rho\left(\vec k\sigma\right)\ln \rho\left(\vec k\sigma\right)\).

\subsection{\(k-\)space mutual information}
Another measure of manybody correlations would be the mutual information between two momentum-space states. For eg., the mutual information \(I_2\) between the states \(\ket{\vec k, \sigma}\) and \(\ket{-\vec k, \sigma}\) is defined as
\begin{equation}\begin{aligned}
	I_2(\vec :-\vec k) = S_\text{EE}(\vec k) + S_\text{EE}(-\vec k) - S_\text{EE}(\vec k,-\vec k)~,
\end{aligned}\end{equation}
where \(S_\text{EE}(\pm\vec k)\) is the von-Neumann entropy associated with the reduced density matrix \(\rho\left(\vec k\sigma\right)\) that has \(\pm \vec k\) traced out, and \(S_\text{EE}(\vec k,-\vec k)\) is associated with the density matrix \(\rho\left(\vec k \sigma, -\vec k \sigma\right)\) that has both \(\vec k\) and \(-\vec k\) traced out. \(\rho\left(\vec k\sigma\right)\) has already been computed in eq.~\ref{density_matrix_ksapce}. In order to calculate \(\rho\left(\vec k \sigma, -\vec k \sigma\right)\), we trace out the configurations of \(-\vec k\) from the density matrix \(\rho\left(\vec k\sigma\right)\):
\begin{equation}\begin{aligned}
	\rho\left(\vec k \sigma, -\vec k \sigma\right) = \sum_{n_{-\vec k\sigma}=0,1} \braket{n_{-\vec k\sigma}| \rho\left(\vec k\sigma\right) | n_{-\vec k\sigma}}
\end{aligned}\end{equation}
By following an approach similar to the one that led to eq.~\ref{overlap_ftransform}, we Fourier transform the \(k-\)space terms:
\begin{equation}\begin{aligned}
	&\rho\left(\vec k \sigma, -\vec k \sigma\right) \\
	&= \sum_{n_{(\vec R_0,\vec R_0 + \vec a,\vec R_0  +\vec a^\prime),\sigma}=0,1}\left[\braket{n_{\vec R_0,\sigma}| \rho\left(\vec k\sigma\right) | n_{\vec R_0, \sigma}} + \xi_{-\vec k}^2\braket{n_{\vec R_0 + \vec a,\sigma}| \rho\left(\vec k\sigma\right) | n_{\vec R_0 + \vec a, \sigma}} + \Xi_{-\vec k}^2\braket{n_{\vec R_0 + \vec a^\prime,\sigma}| \rho\left(\vec k\sigma\right) | n_{\vec R_0 + \vec a^\prime, \sigma}} \right.\\
	&\left.+ 2\xi_{-\vec k}\braket{n_{\vec R_0 + \vec a,\sigma}| \rho\left(\vec k\sigma\right) | n_{\vec R_0, \sigma}} + 2\Xi_{-\vec k}\braket{n_{\vec R_0 + \vec a^\prime,\sigma}| \rho\left(\vec k\sigma\right) | n_{\vec R_0, \sigma}} + 2\xi_{-\vec k}\Xi_{-\vec k}\braket{n_{\vec R_0 + \vec a^\prime,\sigma}| \rho\left(\vec k\sigma\right) | n_{\vec R_0 + \vec a, \sigma}}\right]\\
	&= \sum_{n_{(d,z,1),\sigma}=0,1}\frac{1 + \xi_{\vec k}^2 + \Xi_{\vec k}^2}{3\lambda_{\vec k_0}^2}\left[\braket{n_{d,\sigma} | \rho\left(\vec k\sigma\right) | n_{d,\sigma}} + \frac{1}{\mathcal{W}}\braket{n_{z,\sigma} | \rho\left(\vec k\sigma\right) | n_{z,\sigma}} + \frac{1}{\mathcal{W}}\braket{n_{1,\sigma} | \rho\left(\vec k\sigma\right) | n_{1,\sigma}}\right] \\
				       &+ \sum_{n_{(d,z,1),\sigma}=0,1}\frac{\xi_{\vec k}\left(1 + \Xi_{\vec k}\right)}{2\lambda_{\vec k_0}^2 \mathcal{W}}\left[\braket{n_{d,\sigma} | \rho\left(\vec k\sigma\right) | n_{z,\sigma}} + \braket{n_{z,\sigma} | \rho\left(\vec k\sigma\right) | n_{d,\sigma}} + \braket{n_{z,\sigma} | \rho\left(\vec k\sigma\right) | n_{1,\sigma}} + \braket{n_{1,\sigma} | \rho\left(\vec k\sigma\right) | n_{z,\sigma}} \right]\\
				       &+ \sum_{n_{(d,1),\sigma}=0,1}\frac{\Xi_{\vec k}}{\lambda_{\vec k_0}^2 \mathcal{W}}\left[\braket{n_{1,\sigma} | \rho\left(\vec k\sigma\right) | n_{d,\sigma}} + \braket{n_{d,\sigma} | \rho\left(\vec k\sigma\right) | n_{1,\sigma}}\right]~.
\end{aligned}\end{equation}
